---
title: "mherit Vignette"
author: "Daniel Sprockett"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{mherit Vignette}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Load Pacakages

Load Packages.

```{r load packages, message=FALSE, warning=FALSE}

require("ggplot2")
require("phyloseq")
require("car")
require("mets")
require("mherit")

```

## Import

Import the phyloseq object.

```{r import}

#
ps <- readRDS("/Users/daniel/R_packages/mherit/data/twins_ps.rds")

```

## Subset

Since some twins have mutliple samples, I'll first de-duplicate the subjects. In the future, this may be added to a function as an optional check (i.e. checking to make sure twins' samples were collected on the same day, etc.).

```{r de-duplicate}

samples_to_keep <- get_variable(ps, "Sample_ID")[!duplicated(get_variable(ps, "Subject_ID"))]
ps <- subset_samples(ps, Sample_ID %in% samples_to_keep)

```

For the sake of speeding up this demonstration, next I'll select 400 twin pairs at random to focus on. 

```{r subset}

n <- 400
set.seed(8675309)
families_to_keep <- sample(unique(get_variable(ps, "Family_ID")), n)
ps <- subset_samples(ps, Family_ID %in% families_to_keep)

```

## Filter

Remove taxa not seen at least once in at least 50% of the samples.

```{r filter}

ps <- filter_taxa(ps, function(x) sum(x >= 1) > (0.50*length(x)), TRUE)
ps

```

This leaves us with `r nsamples(ps)` samples and `r ntaxa(ps)` taxa.

Lets check the how many MZ and DZ twins we are left with.

```{r check zygosity}

table(get_variable(ps, "Zygosity"))

```

Ok, not bad. 

## Prepare Dataset

Now I can use the `add_ranks()` function to add higher taxonomic levels to the phyloseq object. For now I'll just add the `Family` and `Genus` levels, but if you leave `ranks` empty, it'll use all of the ranks available in the `tax_table()`.

```{r add_ranks}

ps <- add_ranks(ps, ranks = c("Family", "Genus"))

```

You can see that there are now taxonomic features present in the dataset that are higher than ASVs. 

```{r examine taxa names}

sample(taxa_names(ps), 10)

```

## Normalize

If we take a look at the abundances of the first 2 taxa in the fist 5 samples, we can see this is count data (integers).

```{r pre-normalization}

otu_table(ps)[1:5, 1:2]

```

To make the data more normally distributed, I'll use the `normalize_for_heritability()` function to perform the Box-Cox transformation and then regress on a few covariates in my `sample_data()`. Notice that `Height` is not in the `sample_data()`. The output phyloseq object contains the residuals from the linear regression. In this functions and all of the functions in this package, you can set `verbose = FALSE` to suppress the message. 

```{r normalize_for_heritability}

ps <- normalize_for_heritability(ps, covariates = c("Host_Age", "Host_Sex", "Height"), verbose = TRUE)

```

If you want, you can also add `normalization_plots = TRUE` and the function will output plots for each taxonomic feature showing the distribution before and after normalization. 

Now if we check again, the data is no longer in count form. 

```{r post-normalization}

otu_table(ps)[1:5, 1:2]

```

## Calculate A.C.E.

After I'm satisfied with the normalization, I can use the `calculate_ACE()` function to calculate the A.C.E. model. This function calls the [twinlm](https://www.rdocumentation.org/packages/mets/versions/1.2.3.1/topics/twinlm) function from the `mets` R package. We'll also set `verbose = FALSE` to supress the long list of taxa, like above. 

```{r calculate_ACE}

df <- calculate_ACE(ps, dz = "DZ", zygosity = "Zygosity", family_id = "Family_ID", verbose = FALSE)

```

The output is a data.frame that contains all of the estimates of A, C, and E, as well as their taxonomy.  

```{r}

head(df, 3)

```


## Plot

Finally we can plot the results. 

```{r plot, fig.width=10, fig.height=10}

p <- plot_heritability(df, cutoff = 0.2, color = "Family", verbose = TRUE)
p

```

Don't pay too much attention to the names of the taxa being displayed in this figure. There is a lot of data curation that needs to be done that I am not showing here.

This is a work in progress, so please let me know if you run into any problems. Thank you!
